---
title: "02_eda"
output: html_document
date: "2025-03-15"
---
## Loading required libraries and data
```{r}
library(tidyverse)
#install.packages("reshape")
library(reshape2)

#change the file path to source code location
setwd("C:/Users/Dell/Desktop/UMD/Spring 2025/BUMK 776 Action Learning Project/team8_source_code")
data <- read.csv("data/data_processed.csv")
```

## Visualizing Satisfaction across Cohorts
```{r}
sat_summ <- data %>%
  group_by(MODEL_SCORE) %>%
  summarise(across(c(COVERRECSAT, DATASPEEDSAT, PRICEPLANSAT, BILLPYMTSAT, CUSTSERVSAT, 
                     EASEMGMTSAT, FEATPERKSSAT, PHNUPGDSAT, PHN_TRADESAT, PHN_CREDITSAT, 
                     PHN_FINSAT, PHN_DWNPYMTSAT), ~ mean(.x, na.rm = TRUE)))
sat_summ <- subset(sat_summ, !is.na(MODEL_SCORE)) # remove the row where Model_score=NA

freq_table <- data %>% # add frequency column
  group_by(MODEL_SCORE) %>%
  summarise(Freq = n())

#join the frequency column to the original summary
sat_summ <- left_join(sat_summ, freq_table, by = "MODEL_SCORE")
# Find the frequency column (currently there are freq.y, freq.x, etc)
freq_col <- grep("^Freq", colnames(sat_summ), value = TRUE)

sat_summ <- sat_summ %>%
  rename(Freq = all_of(freq_col)) %>%
  select(MODEL_SCORE, Freq, COVERRECSAT, DATASPEEDSAT, PRICEPLANSAT, BILLPYMTSAT,
         CUSTSERVSAT, EASEMGMTSAT, FEATPERKSSAT, PHNUPGDSAT,
         PHN_TRADESAT, PHN_CREDITSAT, PHN_FINSAT, PHN_DWNPYMTSAT)

#Melt the data and round the mean values to 2 decimal places
sat_summ_heat <- melt(sat_summ[, -2],  
                      id.vars = "MODEL_SCORE", 
                      variable.name = "variable", 
                      value.name = "average_score")

#round the mean_value for visualization
sat_summ_heat$average_score <- round(sat_summ_heat$average_score, 2)
#convert model_score to factor to main order and labels
sat_summ_heat$MODEL_SCORE <- factor(sat_summ_heat$MODEL_SCORE, levels = 1:14)

ggplot(sat_summ_heat, aes(x = variable, y = MODEL_SCORE, fill = average_score)) +
  geom_tile(color = "white") + # Add tile border for clarity
  scale_fill_gradient(low = "white", high = "blue") +
  geom_text(aes(label = average_score), color = "black", size = 3) + # Display values
  theme_minimal() +
  labs(title = "Satisfaction Scores by Cohort", 
       x = "Satisfaction", 
       y = "Cohort") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7)) + # Rotate x-axis labels
  scale_y_discrete(breaks = 1:20) # Use discrete scale for MODEL_SCORE
```
* Cohorts 11, 12 and 13 seem to be consistently dissatisfied relative to other cohorts.

* High satisfaction can be seen for Bill Payment, Customer Serivce and Ease of Management across cohorts.

## Visualizing Likelihood to Continue across Cohorts
```{r}
boxplot(LTC ~ MODEL_SCORE, data=data)

data$lyl_diff <- data$LTC - data$LTR
lyl_summ <- data %>%
  group_by(MODEL_SCORE) %>%
  summarise(across(c(LTC, LTR, lyl_diff), ~ mean(.x, na.rm = TRUE)))
lyl_summ <- subset(lyl_summ, !is.na(MODEL_SCORE))

lyl_summ <- left_join(lyl_summ, freq_table, by = "MODEL_SCORE")
freq_col <- grep("^Freq", colnames(lyl_summ), value = TRUE)

lyl_summ <- lyl_summ %>%
  rename(Freq = all_of(freq_col)) %>%
  select(MODEL_SCORE, Freq, LTC, LTR, lyl_diff)

lyl_summ_heat <- melt(lyl_summ[, -c(2,5)],
                      id.vars = "MODEL_SCORE", 
                      variable.name = "variable", 
                      value.name = "likelihood")

lyl_summ_heat$likelihood <- round(lyl_summ_heat$likelihood, 2) # Round the mean_value for visualization
lyl_summ_heat$MODEL_SCORE <- factor(lyl_summ_heat$MODEL_SCORE, levels = 1:14) #convert model_score to factor to main order and labels

ggplot(lyl_summ_heat, aes(x = variable, y = MODEL_SCORE, fill = likelihood)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "maroon") +
  geom_text(aes(label = likelihood), color = "black", size = 3) +
  theme_minimal() +
  labs(title = "Likelihood by Cohort", 
       x = "Likelihood", 
       y = "Cohort") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7)) +
  scale_y_discrete(breaks = 1:20)
```
* Cohorts 14 and 6 have the highest likelihoods to continue service.

* Low likelihoods can be observed particularly for cohorts 12 and 11.

* Patterns in LTR and LTC seem to be similar across cohorts.

## Potential variables impacting LTC

### Number of Lines
```{r}
data$NUMBER_OF_LINES[data$NUMBER_OF_LINES>=7] <- 7 #this is 7 or more
cor(as.numeric(data$NUMBER_OF_LINES), data$LTC, use = "complete.obs")
data %>%
  group_by(MODEL_SCORE) %>%
  summarise(across(c(NUMBER_OF_LINES,LTC), ~ mean(.x, na.rm = TRUE)))
```
The correlation coefficient shows that more number of lines are not necessarily associated with a higher likelihood to continue.

### Account Tenure
```{r}
cor(data$ACCOUNT_TENURE_DAYS, data$LTC, use = "complete.obs")
data %>%
  group_by(MODEL_SCORE) %>%
  summarise(across(c(ACCOUNT_TENURE_DAYS,LTC), ~ mean(.x, na.rm = TRUE)))
```
The correlation coefficient indicates that longer tenure may be associated with a higher likelihood to continue.

### Word-of-Mouth Sentiment
```{r}
round(prop.table(table(data$MODEL_SCORE, data$WOM_SENTIMENT), margin=1),3)

ggplot(data, aes(x = as.factor(WOM_SENTIMENT), y = LTC)) +
  geom_boxplot(fill = "lightblue") +
  stat_summary(fun = mean, geom = "point", size = 2, color = "red")

data %>%
  group_by(WOM_SENTIMENT) %>%
  summarise(across(LTC, ~ mean(.x, na.rm = TRUE)))
```
Cohorts 11, 12 and 13, who have relatively low LTC scores, show a relatively high proportion of negative WOM sentiments. Negative WOM also translates into significantly lower LTC compared to positive WOM.

### Having an issue recently
```{r}
round(prop.table(table(data$MODEL_SCORE, data$ISS_RECENTISSUE), margin=1),3)
```
The proportion of customers who have had an issue recently is much higher in cohorts that also see a low LTR and a low LTC (cohorts 11, 12 and 13).

### Propensity to Call Care
```{r}
round(prop.table(table(data$MODEL_SCORE, data$HIGH_PROP_TO_CALL_CARE), margin=1),3)
```
The proportion of customers with a high propensity to call care is significantly larger in cohort 12. It is also relatively high in cohorts 7 and 10.

### Primary Reason to Join
```{r}
ggplot(data, aes(x = as.factor(JOI_PRIMARYREASONAP), y = LTC)) +
  geom_boxplot(fill = "lightblue") +
  stat_summary(fun = mean, geom = "point", size = 2, color = "red")
```
While a pattern in the reason for joining and LTC is not very clear, we can see that customers who joined for reasons 8, 9 and 10 seem to have significantly lower LTC scores than other customers.