---
title: "03_model"
output: html_document
date: "2025-03-14"
---

## Loading data
```{r}
#change the file path to source code location
setwd("C:/Users/Dell/Desktop/UMD/Spring 2025/BUMK 776 Action Learning Project/team8_source_code")
data <- read.csv("data/data_processed.csv")

#choosing a subset of variables based on EDA
data <- data[,c('id','LTC','WOM_SENTIMENT','GENDER','MARITALSATUS',
                'KIDSINHH','ETHNICITY','MILITARYSTAT','ACCOUNT_TENURE_DAYS',
                'ACTIVATION_MILESTONE_GROUP_NEW','HIGH_PROP_TO_CALL_CARE','AGE',
                'JOI_PRIMARYREASONAP','NUMBER_OF_LINES','REMAINING_PAYMENTS',
                'MONTHLY_SUBSCRIPTION_COST')]
```

## Converting variables to appropriate formats
```{r}
cat_vars <- c('WOM_SENTIMENT','GENDER','MARITALSATUS','KIDSINHH','ETHNICITY',
              'MILITARYSTAT','ACTIVATION_MILESTONE_GROUP_NEW','AGE')

data[cat_vars] <- lapply(data[cat_vars], as.factor)
data$HIGH_PROP_TO_CALL_CARE <- factor(data$HIGH_PROP_TO_CALL_CARE, levels=c("N", "Y"))
data$WOM_SENTIMENT <- relevel(data$WOM_SENTIMENT, ref="Mostly positive")
data$JOI_PRIMARYREASONAP <- as.factor(data$JOI_PRIMARYREASONAP)
```

##Model 1
```{r}
model1 <- lm(LTC ~ GENDER + MARITALSATUS + KIDSINHH + ETHNICITY + MILITARYSTAT
             + HIGH_PROP_TO_CALL_CARE + AGE + NUMBER_OF_LINES
             + ACCOUNT_TENURE_DAYS + REMAINING_PAYMENTS
             + MONTHLY_SUBSCRIPTION_COST, data = data)
summary(model1)
```

##Model 2
```{r}
model2 <- lm(LTC ~ GENDER + KIDSINHH + MILITARYSTAT + HIGH_PROP_TO_CALL_CARE
             + AGE + NUMBER_OF_LINES + ACCOUNT_TENURE_DAYS + REMAINING_PAYMENTS
             + MONTHLY_SUBSCRIPTION_COST, data = data)
summary(model2)
```

##Model 3
```{r}
model3 <- lm(LTC ~ GENDER + KIDSINHH + MILITARYSTAT + HIGH_PROP_TO_CALL_CARE
             + AGE + NUMBER_OF_LINES + ACCOUNT_TENURE_DAYS + REMAINING_PAYMENTS
             + MONTHLY_SUBSCRIPTION_COST + JOI_PRIMARYREASONAP, data = data)
summary(model3)
```

##Final Model
```{r}
final_model <- lm(LTC ~ AGE + HIGH_PROP_TO_CALL_CARE + MONTHLY_SUBSCRIPTION_COST
                  + MILITARYSTAT + ACCOUNT_TENURE_DAYS + JOI_PRIMARYREASONAP
                  + WOM_SENTIMENT, data = data)
summary(final_model)
```

## Summary
We iteratively refined our model by testing different variable combinations removing less significant predictors, and introducing key explanatory factors to enhance predictive power.

The final model retains only the most impactful variables, achieving a significantly improved fit with an adjusted R-squared of 0.6387.

## Proportion of Word-Of-Mouth Sentiments across cohorts
```{r}
data$sentiment <- as.character(data$WOM_SENTIMENT)
data$sentiment[!is.na(data$sentiment) & (data$sentiment=="Mostly positive" | data$sentiment=="All positive")] <- "Positive"
data$sentiment[!is.na(data$sentiment) & (data$sentiment=="Mostly negative" | data$sentiment=="All negative")] <- "Negative"

prop.table(table(data$sentiment, data$MODEL_SCORE), margin=2)
```